Unsupervised embedding of trajectories captures the latent structure of mobility

### 2020-10-19 paper/current/main.tex
one more thing on first main para
some minor issues

We show that the embedding distance better predicts actual mobility flows than the geographic distance.
In the case of flight itineraries, the embedding distance explains more than twice the expected flux between airports ($R^{2} = 0.51$, Fig.~\ref{fig:gravity_performance}a) than does geographic distance ($R^{2} = 0.22$).  Also, the embedding distance produces better predictions of actual flux between airports than does the geographic distance (Fig.~\ref{fig:gravity_performance}b).
In the case of South Korean hotel accommodations, such that the embedding distance better explains the expected flux ($R^{2} = 0.57$, Fig.~\ref{fig:gravity_performance}c) than does geographic distance ($R^{2} = 0.25$), and predictions made using embedding distance outperform those made with geographic distance (Fig.~\ref{fig:gravity_performance}d). 
This performance is consistent in the case of scientific mobility:
the embedding distance explains more than twice the expected flux ($R^{2} = 0.48$, Fig.~\ref{fig:gravity_performance}e) than does the geographic distance ($R^{2} = 0.22$), and predictions made using the embedding distance outperform those using the geographic distance (Fig.~\ref{fig:gravity_performance}f).
These patterns hold for the subsets of only domestic (within-country organization pairs, Fig.~\ref{fig:supp:gravity_correlation}c and Fig.~\ref{fig:supp:predict_geo_bymodel}) and only international mobility flows (across-country organization pairs, Fig.~\ref{fig:supp:gravity_correlation}d and  Fig.~\ref{fig:supp:predict_emb_bymodel}).
The embedding distance also out-performs alternative diffusion-based network distance measures including the personalized-page rank scores  calculated from the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}).
In sum, our results demonstrate that the embedding distance better captures patterns of actual mobility than does the geographic distance consistently.

@@

We report that this curve is almost universal across many countries. 
For instance, China's curve closely mirrors that of the United States (Fig.~\ref{fig:length}b).
Smaller but scientifically advanced countries such as Australia, and other populous countries such as Brazil also exhibit curves similar to the United States (Fig.~\ref{fig:length}b, inset), though their exact position and size vary.
Other nations exhibit different curves, such as Egypt which has an``S''-shaped curve, meaning that their largest organizations are not as well connected as other countries' largest organizations. 
Similar patterns can be found across many of the 30 of the countries with the most total researchers (Fig.~\ref{fig:concave30}) see \SI{} for more discussion).

@@

Processed data, analysis, and associated code have been made publicly available and can be found at \url{https://github.com/murrayds/sci-mobility-emb}.

@@

%
% Author Contributions
% 
\subsection*{Author Contributions}
Dakota Murray and Jisung Yoon were involved in all stages of data handling, analysis, and writing
Sadamori Kojaku contributed to the development of the mathematical framework used in this paper.
Rodrigo Costas provided access to and knowledge of the dataset of scientific mobility.
Woo-Sung Jung, Staša Milojević , and Yong-Yeol Ahn provided guidance, mentorship, and conceptualization.
All authors contributed to the preparation of this manuscript.

%
% Supporting Information
% 
\subsection*{Additional Information}
Supplementary Information is available for this paper.
Correspondence and requests for materials should be addressed to Dr. Staša Milojević and Dr. Yong-Yeol Ahn”.
Reprints and permissions information is available at www.nature.com/reprints”

>>>>

We show that the embedding distance better predicts actual mobility flows than the geographic distance across three disparate datasets.
In the case of flight itineraries, the embedding distance explains more than twice the expected flux between airports ($R^{2} = 0.51$, Fig.~\ref{fig:gravity_performance}a) than does geographic distance ($R^{2} = 0.22$).  Also, the embedding distance produces better predictions of actual flux between airports than does the geographic distance (Fig.~\ref{fig:gravity_performance}b).
In the case of South Korean hotel accommodations, such that the embedding distance better explains the expected flux ($R^{2} = 0.57$, Fig.~\ref{fig:gravity_performance}c) than does geographic distance ($R^{2} = 0.25$), and predictions made using embedding distance outperform those made with geographic distance (Fig.~\ref{fig:gravity_performance}d). 
This performance is consistent in the case of scientific mobility:
the embedding distance explains more than twice the expected flux ($R^{2} = 0.48$, Fig.~\ref{fig:gravity_performance}e) than does the geographic distance ($R^{2} = 0.22$), and predictions made using the embedding distance outperform those using the geographic distance (Fig.~\ref{fig:gravity_performance}f).
These patterns hold for the subsets of only domestic (within-country organization pairs, Fig.~\ref{fig:supp:gravity_correlation}c and Fig.~\ref{fig:supp:predict_geo_bymodel}) and only international mobility flows (across-country organization pairs, Fig.~\ref{fig:supp:gravity_correlation}d and  Fig.~\ref{fig:supp:predict_emb_bymodel}).
The embedding distance also out-performs alternative diffusion-based network distance measures including the personalized-page rank scores  calculated from the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}).
In sum, our results demonstrate that the embedding distance better captures patterns of actual mobility than does the geographic distance consistently.

@@

We report that this curve is almost universal across many countries. 
For instance, China's curve closely mirrors that of the United States (Fig.~\ref{fig:length}b).
Smaller but scientifically advanced countries such as Australia, and other populous countries such as Brazil also exhibit curves similar to the United States (Fig.~\ref{fig:length}b, inset).
Other nations exhibit different curves which lacks the portions with decreasing norm, probably indicating the lack of highly prestigious institutions.
Similar patterns can be found across many of the 30 of the countries with the most total researchers (Fig.~\ref{fig:concave30}) see \SI{} for more discussion).

@@

@@

%
% Author Contributions
% 
\subsection*{Author Contributions}
Dakota Murray and Jisung Yoon were involved in all stages of data handling, analysis, and writing.
Sadamori Kojaku contributed to the development of the mathematical framework used in this paper.
Rodrigo Costas provided access to and knowledge of the dataset of scientific mobility.
Woo-Sung Jung, Staša Milojević, and Yong-Yeol Ahn provided guidance, mentorship, and conceptualization.
All authors contributed to the preparation of this manuscript.

%
% Supporting Information
% 
\subsection*{Additional Information}
Supplementary Information is available for this paper.
Correspondence and requests for materials should be addressed to Dr. Staša Milojević and Dr. Yong-Yeol Ahn.
Reprints and permissions information is available at www.nature.com/reprints
